---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6,
  fig.width = 8
)
rstan::rstan_options(silent = TRUE, open_progress = FALSE,
                     show_messages = FALSE)

if (! "bayesplot" %in% installed.packages()){
  install.packages("bayesplot")
}
```
Somatic growth of fishes is a fundamental trait that determines essential ecosystem services such as food provision and nutrient cycling. 
Growth rate information can be derived through age estimation based on the analysis of sagittal otoliths. While fitting growth models on size-at-age data is the most frequently employed approach to deriving growth parameters, this method requires a high number of individuals. 
An alternative approach based on back-calculation can provide approximations to individual-level growth trajectories. 
`fishgrowbot` provides functions to perform the back-calculation in a Bayesian framework. 
Further, the package provides a Bayesian framework to fit the von Bertalanffy growth model to the back-calculated lengths in a hierarchical structure. 
Finally, `fishgrowbot` provides functions to visualize the results. 
In this vignette, we provide an introduction to `fishgrowbot` through a case study. For more theoretical background of the methodology, see [Morat et al. (2020)](https://doi.org/10.1038/s41597-020-00711-y)         


First things first: we need to load some packages. 

```{r setup}
library(fishgrowbot)
library(dplyr)
```

To introduce the functionalities of `fishgrowbot`, we look at an example for
*Epinephelus merra*. The function `bcalc()` returns both a dataframe with the
back-calculated lengths and their uncertainty and the model object for more
details on the fit of the `bc` stan model.

The input data should contain:         
  + id: Unique fish id per individual.  
  + radi: Measurements of otolith growth rings (in mm).  
  + agei: Age estimation of fish.  
  + lencap: Length at capture (in mm).  
  + radcap: Radius of otolith at capture (in mm).  
  + l0p: Length of fish at hatching (in mm).  

```{r, bc, message = FALSE}
# get data
em <- filter(coral_reef_fishes_data, species == "Epinephelus merra",
             location == "Moorea")
# back-calculation
bc <- bcalc(data = em)
head(bc$lengths)
```

The function `bcplot` helps visualize the back calculation. Some examples:

```{r, plot1a, message = FALSE}
bcplot(bc$lengths)
```

```{r, plot1b, message = FALSE}
bcplot(bc$lengths, colorid = TRUE)
```

```{r, plot1c, message = FALSE}
bcplot(bc$lengths, colorid = TRUE, facet = TRUE, error = TRUE)
```

Then, we can fit the hierarchical von Bertalanffy growth model that allows for
the estimation of multiple parameters. Importantly, length measures should be
given in cm.

```{r, growthmodel, message = FALSE}
# fit growth model
growthmodel <- growthreg(length = bc$lengths$l_m / 10, age =  bc$lengths$age,
                         id =  bc$lengths$id, lmax = 32, linf_m = 28,
                         linf_sd = 5, l0_m = 0.15, l0_sd = 0.015, iter = 4000,
                         open_progress = FALSE, plot = FALSE)
# summary growth parameters
growthmodel$summary
```
We can use existing packages `rstan` and `bayesplot` to make diagnostic plots. 

```{r}
library(rstan)
library(bayesplot)
# Posterior predictive check
yrep <- extract(growthmodel$stanfit, "y_rep")$y_rep
ppc_dens_overlay(y = bc$lengths$l_m / 10, yrep = yrep[1:100,])

#trace plots
posterior <- extract(growthmodel$stanfit, permuted = FALSE)

p <- mcmc_trace(posterior,  pars = c("linf", "k"),
                facet_args = list(nrow = 2, labeller = label_parsed))
p + facet_text(size = 15)
```




Now we can visualize the fit with the function `gmplot()`.

```{r, plot2a, message = FALSE}
gmplot(growthmodel)
```

Setting `id = TRUE`, we can see the individual growth curves.

```{r, plot2b, message = FALSE}
gmplot(growthmodel, id = TRUE)
```

If there are many individuals, the id graph can quickly get crowded. Set
`facet = TRUE` to have a subplot per individual.

```{r, plot2c, message = FALSE}
gmplot(growthmodel, id = TRUE, facet = TRUE)
```
