---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.height = 6, 
  fig.width = 8
)
rstan::rstan_options(silent = TRUE, open_progress=FALSE,show_messages=FALSE)
```

First things first: we need to load some packages. 
```{r setup}
library(fishgrowbot)
library(dplyr)
```

To introduce the functionalities of `fishgrowbot`, we look at an example for *Epinephelus merra*. 
The function `bcalc()` returns both a dataframe with the back-calculated lengths and their uncertainty and the model object for more details on the fit of the bc stan model. 

```{r, bc, message = FALSE}
# get data
em <- dplyr::filter(fishgrowbot::coral_reef_fishes_data, species == "Epinephelus merra", location == "Moorea")

# back-calculation
bc <- fishgrowbot::bcalc(data = em)
head(bc$lengths)
```

The function `bcplot` helps visualize the back calculation. Some examples:

```{r, plot1a, message = FALSE}
bcplot(bc$lengths)
```

```{r, plot1b, message = FALSE}
bcplot(bc$lengths, colorid = TRUE)
```

```{r, plot1c, message = FALSE}
bcplot(bc$lengths, colorid = TRUE, facet = TRUE, error = TRUE)
```

Then, we can fit the hierarchical Von Bertalanffy growthmodel that allows for the estimation of multiple parameters.     

```{r, growthmodel, message = FALSE}
# fit growth model
growthmodel <- fishgrowbot::growthreg(length = bc$lengths$l_m/10, 
                                      age =  bc$lengths$age,
                                      id =  bc$lengths$id, 
                                      lmax = 32, 
                                      linf_m = 28, 
                                      linf_sd = 5, 
                                      l0_m = 0.15, 
                                      l0_sd = 0.015, 
                                      iter = 4000,
                                      open_progress = FALSE,
                                      plot = FALSE)

# summary growth parameters
growthmodel$summary
```


Now we can visualize the fit with the function `gmplot()`.

```{r, plot2a, message = FALSE}
gmplot(growthmodel)
```

Setting id to TRUE, we can see the individual growth curves.

```{r, plot2b, message = FALSE}
gmplot(growthmodel, id = TRUE)
```

If there are many individuals, the id graph can quickly get crowded. Set facet = TRUE to have a subplot per individual. 

```{r, plot2c, message = FALSE}
gmplot(growthmodel, id = TRUE, facet = TRUE)
```



