---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6,
  fig.width = 8
)
rstan::rstan_options(silent = TRUE, open_progress = FALSE,
                     show_messages = FALSE)
```
In this vignette, we provide an introduction to `fishgrowbot` through a case study. For theoretical background of the methodology, see [Morat et al. (2020)](https://doi.org/10.1038/s41597-020-00711-y)         


First things first: we need to load some packages. 

```{r setup}
library(fishgrowbot)
library(dplyr)
```

To introduce the functionalities of `fishgrowbot`, we look at an example for
*Epinephelus merra*. The function `bcalc()` returns both a dataframe with the
back-calculated lengths and their uncertainty and the model object for more
details on the fit of the `bc` stan model.

The input data should contain:
  + id: Unique fish id per individual.  
  + radi: Measurements of otolith growth rings (in mm).  
  + agei: Age estimation of fish.  
  + lencap: Length at capture (in mm).  
  + radcap: Radius of otolith at capture (in mm).  
  + l0p: Length of fish at hatching (in mm).  

```{r, bc, message = FALSE}
# get data
em <- filter(coral_reef_fishes_data, species == "Epinephelus merra",
             location == "Moorea")
# back-calculation
bc <- bcalc(data = em)
head(bc$lengths)
```

The function `bcplot` helps visualize the back calculation. Some examples:

```{r, plot1a, message = FALSE}
bcplot(bc$lengths)
```

```{r, plot1b, message = FALSE}
bcplot(bc$lengths, colorid = TRUE)
```

```{r, plot1c, message = FALSE}
bcplot(bc$lengths, colorid = TRUE, facet = TRUE, error = TRUE)
```

Then, we can fit the hierarchical von Bertalanffy growth model that allows for
the estimation of multiple parameters. Importantly, length measures should be
given in cm.

```{r, growthmodel, message = FALSE}
# fit growth model
growthmodel <- growthreg(length = bc$lengths$l_m / 10, age =  bc$lengths$age,
                         id =  bc$lengths$id, lmax = 32, linf_m = 28,
                         linf_sd = 5, l0_m = 0.15, l0_sd = 0.015, iter = 4000,
                         open_progress = FALSE, plot = FALSE)
# summary growth parameters
growthmodel$summary
```

Now we can visualize the fit with the function `gmplot()`.

```{r, plot2a, message = FALSE}
gmplot(growthmodel)
```

Setting `id = TRUE`, we can see the individual growth curves.

```{r, plot2b, message = FALSE}
gmplot(growthmodel, id = TRUE)
```

If there are many individuals, the id graph can quickly get crowded. Set
`facet = TRUE` to have a subplot per individual.

```{r, plot2c, message = FALSE}
gmplot(growthmodel, id = TRUE, facet = TRUE)
```
